Dynamic capabilities and competitive advantage

### 2019-12-11 article4_jbr.rnw
Ready for submission

\title{A Tale of Two Mechanisms: How mediators act and interact in dynamic capabilities
  theory\\ Lock, Stock and Two Convoking Mechanisms: Towards a holistic theory of dynamic
  capabilitie }
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

@@

\begin{abstract}
This paper investigates strategic cognition as mediating factor in the dynamic
capabilities-competitive advantage relationship. By including strategic cognition into the
analysis in addition to the conventional mediator (operating capabilities) we argue that a
more holistic approach to understanding the workings of dynamic capabilities emerges. We
introduce two expressions of how dynamic capabilities work, namely as a conductor (the
conventional mechanism) and as a coach (the cognitive mechanism). Using repeated
cross-sectional data from Norwegian technology firms our results indicate support for our
holistic model. This has implications for understanding how dynamic capabilites work, as
well as practical implications for how strategic change is implemented in firms.

@@

% paper about the DC for MNC in india an dimpact on CA
%https://journals.sagepub.com/doi/abs/10.1177/0971890717701781

Extant literature on dynamic capabilities are increasingly focusing on how firm resources
are changed by dynamic capabilities as a response to changes in the environment
\citep{Schilke2018,Eriksson2014}, thus placing firm capabilities as a mediator between
dynamic capabilities and competetive advantage \citep{Helfat2007}. This notion of
'orchestration' of firm resources, assets and capabilities has indeed been emphasized in
the extant literature \citep{Sirmon2011,Helfat2015}. Less emphasis has been placed on how
other mechanisms can be work in shaping firm competitive advantage. Indeed, this gap in
the literature is explicitly emphasized by \cite{Schilke2018} where they 'see an
interesting opportunity for future work to add greater richness to our understanding of
the mechanism of resource-base change, given this mechanism’s central role in many
foundational works (..) and the diverse ways in which resource changes can potentially
come about' (p. 419).

The conventional mechanism explored in the extant literature is related to changes in the
firm`s resource base \citep{Eriksson2014,Schilke2018}. This mechanism is also included in
the dominating definition of dynamic capabilities found in \citep{Helfat2007}: 'A dynamic
capability is the capacity of an organization to purposefully create, extend or modify its
resource base' (p. 4). This definition opens up for a variety of components
included in the term 'resource base', a point we discuss later in this paper. We coin this
expression of dynamic capabilities as the {\bf conductor} envisioning a conductor
orchestrating firm capabilities.

This notion of a {\bf conductor} albeit very important to the development of dynamic
capabilities literature, seems to leave out a considerable part of a firm`s 'resource
base', namely those resources that are tied to how the employees of the firm sees their
mission, internalize information and effectuates change. These traits of the organization
is very much related to concepts of 'mindset' \citep{Dweck2016}, 'employee cognition'
\citep{Lakoff1987,Gavetti2012} 'organizational genes' \citep{Nelson1982} and
'sense-making' \citep{Weick1995} to name but a few. However, in relations with dynamic
capabilities, one promising avenue for exploring mechanisms, beyond the conventional ones
described above, makes a clear distinction between \emph{behavioral} and
\emph{non-behavioral} objects of change \citep{Verona2011}. These non-behavioral objects
of change contain psychological objects such as cognitive frames, motivation and identity
that are important for strategy formation and implementation \citep{Vince2011,Gavetti2012}
as they 'concerns how decisions or actions are shaped, subverted and transformed by
emotions, and (..) determine the "way we do things here"' \citep[p. 338]{Vince2011}. In
this respect they work as firm resources in generating outcomes such as performance and
competitive advantage. This notion is popularly perhaps expressed most clearly by Peter
Drucker`s famous quote about how 'culture eats strategy for lunch'. \footnote{The debate
  on whether Drucker actually uttered these words are still in the open, but it was
  attributed to him by Mark Fields in 2006} Moreover, these non-behavioral objects of
change are themselves far from static and are found to change over time as responses to
external change \citep{Zollo2016} thus making them a natural part of the mechanisms
underpinning how dynamic capabilities work \citep{Schilke2018,Verona2011}.  Still, far less
is know about the non-behavioral objects of change and how they work as a mechanism in the
dynamic capabilities theory.We coin {\bf coach} to capture the enabling role dynamic
capabilities plays through \emph{strategic cognition} as
juxtaposed to the more directional role played as {\bf conductor} of \emph{operational capabilities}.

We aim to contribute to bridging this gap by theorize about how non-behavioral objects of
change work as a mechanism through which dynamic capabilities influence competitive
advantage. Specifically, we ask the following research question:

@@

To shed light on this important question we theorize on the role of a particular type of
non-behavioral object of change, \emph{strategic cognition}, and how it acts as a mediator
between dynamic capabilities and firm competitive advantage and hence has a pivotal place
in the literature. We build on important work on the concept of organizational cognition
\citep{Lakoff1987} , non-behavioral object of change \citep{Verona2011} and cognitive
psychology \cite{Tversky1983,Kahneman2011} and hypothesize about how strategic cognition
acts as a mediator in the same way as conventional behavioral objects of change (such as
routines and capabilities) and that these two are complementary to a certain
degree. Consequently, we argue, dynamic capabilities takes the function of a {\bf conductor}
by orchestrating routines and capabilities, and the role of a {\bf coach} in stimulating
strategic cognition. We use the case of Microsoft's transition into the cloud computing
business as an illuminating case for demonstrating how the roles of {\bf conductor} and
{\bf coach} coexist and act partly complimentary in generating strategic change and
competitive advantage.

Centering this paper around the dynamic capabilities-competitive advantage nexus stems
from the origins and promises of the dynamic capabilities theory. Competitive advantage is
considered the 'holy grail' of strategic management \citep{Schilke2018} and is indeed the
focal point of important antecedent theories of dynamic capabilities
\citep{Arndt2018}. Consequently, contributions to push the theory forward should enable a
better understanding of \emph{how} dynamic capabilities work to influence firms'
competitive advantage. In this spirit, our paper makes the following contributions to
extant dynamic capabilities theory. First, we provide clarification on the theoretical
role of non-behavioral objects of change as well as a possible empirical
operationalization through the concept of \emph{proactiveness}. Second, we conceptualize
two different expressions of dynamic capabilities (i.e. {\bf conductor} and {\bf coach})
and theorize on how they work in the dynamic capabilities-competitive advantage nexus. We
also extend this reasoning to include how they relate to each other. Finally, we develop and
test an empirical model suited for analyzing mediating constructs on the dynamic
capabilities-competitive advantage nexus. The time-lag of our data in conjunction with
advances in empirical estimations of multiple simultaneous mediators makes for an
interesting model to push our understanding of important mechanisms in the
theory.

@@

\section{Definitions}
The collective term for assets available for firms in their pursuit of their objectives
are \emph{resources}. This term entails tangible assets such as machines, capital, and
labor, but also intangible assets such as procedures, practices, and intellectual
property. The process of using these resources in the pursuit of business objectives such
as competitive advantage is a firm's \emph{capabilities} which can be a set of routines
and skills applied. Resources and capabilities are both tied organizational behavior and
how the organization act in their operation. Hence, they are coined \emph{behavioral
  objects of change} \citep{Verona2011}: 'A pattern of action, a process, an operating
routine, or any form of group activity characterized by some level of stability and
predictability' (ibid. p.538). Such behavioral objects such as capabilities are the
conventional understanding of how dynamic capabilities work; they 'purposefully create,
extend and modify the resource base' (helfat). In the terminology proposed in this paper
this process takes the shape as a 'conductor'.

However, as \cite{Verona2011} points out: 'This approach misses the fundamental aspect
that organizational capabilities specific to managing change can hardly be reduced to the
management of behavioral change' (ibid p. 538). This opens for \emph{non-behavioral
  objects of change} which are defined as 'behavioral antecedents, such as cognition
(e.g. cognitive frames) and motivation' (ibid p. 539). This concept entails the 'how
decisions or actions are shaped, subverted, and/or transformed by emotions; and it
concerns how emotions become embedded in cultural and political practices that determine
the "way we do things here"' (Vince and Gabriel 2011 . ch 15). One particular dimension of
these non-behavioral objects of change particularly conducive to the context of
generating strategic change is the \emph{strategic cognition} of the
organization. Strategic cognition can be defined as conceptual structures in the minds of
the individuals that encapsulate a shared understanding of the reality the
individuals face with respect to the strategic goal of the organization \footnote{This
  definition is based on the definition of \cite{Lakoff1987} definition of the more generic
  'cognitive representation' and is adapted to fit the pursuit of strategic goals such as
  competetive advantage which is the aim of this paper}. This strategic
cognition is the result of a mental process that underlie internal
audiences adoption of new strategic representation and identity code to move toward a
new or existing goal \citep{Gavetti2012}. Although strategic cognition may lead to
behavioral outcome, the nature of the concept itself is tied to cognition rather than
actual latent action \citep{Zollo2016}. Thus, dynamic capabilities are not sufficiently
creating strategic change through its role as 'conductor'. Additionally they need to
exercise a role of 'coach' to adapt the strategic cognition of the organization to create
a shared understanding of the sense and necessity of the strategic goals of the
orientation.

\section{Hypotheses}\label{sec:hyp}

%DC as resource orchestration - influencing capabilitie
%DC as cognitive programmer - inluencing cognitive frames

%Prior research focused on managerial cognition, but gavetti shows that hierarchy matters
%and different cognition maters

What doe sit mean to execute DC (sense etc) and see opportunities and change underlying...

@@

The non-behavioral objects of action has no similar agreed upon analogy, so for the
purpose of our analysis we envision this mechanism to work more as a cognitive process
enabling the organization to tear away from their current understanding and ways in which
things are done \citep{Vince2011}. Whereas the analogy of a {\bf conductor} suggests a
function that composes an outcome based on the available capabilities, the notion of a
{\bf conductor} refers to the function of enabling others to achieve their objectives. The process
is all about changing an organization`s thinking from the automated and well-known, to the
more cognitively aware, moving from what Kahneman and Tversky coins the automated 'system
one' to the deliberate 'system two' thinking \citep{Kahneman2011}. We will now turn to the two functions
of dynamic capabilities in turn.

@@

The bulk of prior contributions study firm capabilities and resources as the mechanism
through which dynamic capabilities influence an outcome
\citep{Fainshmidt2016a,Protogerou2012,Schilke2018,Eriksson2014}. Particularly, much of the
extant literature concerns measures of accounting profitability and competitive advantage
as the final outcome of dynamic capabilities
\cite{shamsie2009,Schilke2014,Schilke2014a,Teece2016}. Thus, from these prior
contributions we derive a fundamental hypotheses of the relationship between dynamic
capabilities and competitive advantage:

@@

Juxtaposed to the experiences from Microsoft, the case of Polaroid brings another set of
experiences to our discussion on non-behavioral objects of action (see \citep{Tripsas2000}
for a detailed account). Cognitive inertia on the part of the organization played a
pivotal part in the failing of the company`s digital imaging effort \citep{Verona2011}. In
other words, Polaroid did not exercise dynamic capabilities to change the cognition of the
organization. Similar patterns are found in Kodak and Anderson Consulting
\citep{Kaplan2005}, and Smith Corona \citep{Danneels2010}. These examples highlight the
absence of 'dynamic capabilities specific to the adaptation of cognitive frames, first,
and consequently of operating processes' \citep[p. 541]{Verona2011}. This is also
highlighted at the core of the dynamic capabilities literature: 'Dynamic capabilities are
about doing the right things, at the right time, based on new product (and process)
development, unique managerial orchestration processes, a strong and {\bf change-oriented
  organizational culture}' (our emphasis) \cite[p]{Teece2014}.

When dynamic capabilities are executed it is as a response to changes in the environment
in order to stay competitive. This processes relates to creating strategic change
\citep{Helfat2007} and entails changing strategic direction and objectives of the
organization. This strategic change requires, in addition to modifying, extending and
creating firm capabilities, also some kind of alignment with the organization as a whole -
what is often referred to as 'how we do things here' \citep{Nelson1982}, and how an
organization thinks and act \citep{Nadella2017}. This alignment is possible only if the
employees share a common cognitive frame related to the the strategic direction of the
company enabling them to make sense of the changes ahead. This is what we have coined
\emph{strategic cognition} and has its expression in Microsoft's development of a growth
mindset. Strategic cognition is needed to get employees on board as well as capture the
efficiency of non-directed change. Getting on board refers to the general willingness of
the organization to accept and value the importance of some decision or system 'without
being prompted or required by formal governance
mechanisms'\citep[p. 812]{Simsek2009}. Moreover, When individuals of a firm buys into
changes they will enhance the full organization`s ability to not only act upon changes
\citep{Marchand2004}, but also in 'timely responsiveness' \citep{Kohli1993}. Consequently,
strategic cognition not only contributes to getting people on board, it also maintain a
more efficient change process where the individuals take more responsibilities for the
change itself and leverages their individual knowledge in the endeavor. For a firm to
create real strategic change as enabled by their dynamic capabilities and hence stay
competitive, is will need to act as a {\bf coach} to instill a strategic cognition in the
organization.

@@

\subsection{Proactiveness as strategic cognition}

@@

We collected survey data from R\&D active firms in Norway at two points in time
($T_0=2005$ and $T_1=2014$). The population was all businesses registered to a scheme for
tax deduction of R\&D costs (called SkatteFUNN). As all enterprises which are eligible for
taxation could register their R\&D activities to receive a tax refund, the registered
enterprises include close to all enterprises which are involved in such activities at the
time of our study. The context of R\&D active firms is particularly relevant for studying
dynamic capabilities because R\&D activity is a signal of responsiveness to change and
that the firms are located in industries with a certain level of dynamism.

We contacted all the firms that were registering R\&D activities during May to December
2005 yielding a total of 1721 firms. Using a web-based questionnaire we included items to
measure the key constructs of this paper. Each firm received an email with a link to the
questionnaire within a month after their registration for the R\&D tax credit. This
initial contact was followed by two subsequent email reminders. A total of 1199 (70 \% of
the initial sample size) answered and returned the questionnaire. In the second phase of
the study each of these 1199 were contacted again in 2014, 9 years later, with the lion
share of the firms contacted during the spring and summer of 2014. All received a similar
web-based questionnaire containing the same measures of the focal constructs of this
paper. 262 of the firms returned filled-in questionnaire.

@@

\section{Results}
The main results of our analysis are depicted both in the graphical path diagrams (figure
\ref{fig:res}) and the regression table in table \ref{tab:res}. The use of PLS-SEM has
increased substantially in the realm of marketing \citep{hult2018} and strategic management
\citep{Hair2012} and is the topic of special issues in leading journals (e.g. Journal of
Business Research call for papers in 2020). In this tradition we aim to present thorough
analysis of our models and largely follows the recommendations made in recent
contributions \citep{Hair2019}.

@@

\section{Discussion}
The aim of this paper is to theorize about how non-behavioral objects of change work as a
mechanism through which dynamic capabilities influence competitive advantage. Our results
indicate that both behavioral (\emph{operating capabilities}) and non-behavioral
(\emph{strategic cognition}) mediates the effect from dynamic capabilities on competitive
advantage. The first mechanism, dubbed 'conducting' in reference to the well established
terminology of 'resource orchestration' \citep{Sirmon2011,Helfat2007}, shows a robust and
strong function as a mediator in line with our first hypothesis. When firms exhibit strong
dynamic capabilities they are able to adapt their operating capabilites to stay
competitive. We hypothesized that this relationship would be positive and significant. Our
results indicated exactly this relationship and thus lending support to hour hypothesis.
In our example context of Microsoft Nadella did several investments in capabilities and
resources as a response to Microsofts dynamic capabilities execution. For example, he
acquired important capabilities from M\&A efforts, and upgraded the competencies regarding
cloud business and big data \citep{Nadella2017}. This should not be surprising given the vast
literature on indirect effects of dynamic capabilities \citep{Eriksson2014} supported by
empirical and theoretical research
\citep{Fainshmidt2016a,Protogerou2012,Karimi2015}. However, it adds to this ongoing effort
of investigating mechanisms through which dynamic capabilities work which is called for in
the literature \citep{Schilke2018}.

>>>>

\title{One goal, two paths: How dynamic capabilities enable competitive advantage through behavioral and non-behavioral objects of change}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

@@

\begin{abstract}
% This paper investigates strategic cognition as mediating factor in the dynamic
% capabilities-competitive advantage relationship. By including strategic cognition into the
% analysis in addition to the conventional mediator (operating capabilities) we argue that a
% more holistic approach to understanding the workings of dynamic capabilities emerges. We
% introduce two expressions of how dynamic capabilities work, namely as a conductor (the
% conventional mechanism) and as a coach (the cognitive mechanism). Using repeated
% cross-sectional data from Norwegian technology firms our results indicate support for our
% holistic model. This has implications for understanding how dynamic capabilites work, as
% well as practical implications for how strategic change is implemented in firms.

  The ultimate goal in the theory of dynamic capabilities (DCs) is competitive advantage.
  This paper analyzes to what extent DCs influence competitive advantage through two paths
  - one old, and one new - namely behavioral and non-behavioral objects of change.  In
  addition to emergent theorizing and real world examples, we use the metaphor of “DC as
  coach” and DC as conductor” to build a conceptual model and derive hypotheses.  Using
  longitudinal survey data with a 10 year time lag we find that DCs influence competitive
  advantage through both paths. Thus, the paper empirically validates emergent theorizing
  proposing non-behavioral objects of change as a new path between DC and competitive
  advantage. Moreover, the paper helps to broaden the conceptual basis for empirically
  investigating how DCs enable competitive advantage. This has implications for
  understanding how dynamic capabilities work, as well as practical implications for how
  strategic change is implemented in firms.

@@

% paper about the DC for MNC in india an dimpact on CA
%https://journals.sagepub.com/doi/abs/10.1177/0971890717701781
The ultimate goal in the theory of dynamic capabilities ($DC$) is competitive advantage
\citep{Schilke2014,Helfat2007,Protogerou2012,Li2014,Efrat2018a,Mikalef2017,Davcik2016a}. Extant
theory argues that competitive advantage ($CA$) stem from resources and routines that are
well aligned to firms’ competitive environment
\citep{Helfat2007,Peteraf2013,Teece1997,Li2014,Jantunen201,Makkonen2014}. Dynamic
capabilities are a source of longer-term competitive advantage in this perspective through
its capacity to purposefully change these lower-level routines and capabilities
\citep{Winter2003,Helfat2011,Schilke2018,Pezeshkan2016b,Lin2014a}, also referred to as
ordinary, substantive, and zero-level routines and capabilities
\citep{Winter2003,Danneels2008,Collis1994}. As a group, these lower level routines and
capabilities are called 'behavioral objects of change' \citep{Verona2011,Zollo2016}, and constitutes an
important path through which DCs can enable competitive advantage \citep{Verona2011}.

However, less emphasis has been placed on how other mechanisms can be work in shaping firm
competitive advantage. Indeed, this gap in the literature is explicitly emphasized by
\citep{Schilke2018} where they 'see an interesting opportunity for future work to add
greater richness to our understanding of the mechanism of resource-base change, given this
mechanisms central role in many foundational works (..) and the diverse ways in which
resource changes can potentially come the mechanism of resource-base change, given this
mechanisms central role in many foundational works (..) and the diverse ways in which
resource changes can potentially come about' (p. 419).

Indeed, the extant conceptualization of the $DC-CA$ link leaves out a considerable part of
a firm‘s 'resource base' that are non-behavioral, such as resources that are tied to how
the employees of the firm sees their mission, internalize information and effectuates
change. These traits of the organization is very much related to concepts of 'mindset'
\citep{Dweck2016}, ’employee cognition’ \citep{Lakoff1987,Gavetti2012} 'organizational
genes' \citep{Nelson1982} and 'sense-making' \citep{Weick1995} to name but a few. However,
in relations with dynamic capabilities, one promising avenue for exploring mechanisms,
beyond the conventional behavioral objects of change is to examine the role of
\emph{non-behavioral} objects of change in the theory of DC \citep{Verona2011}. These
non-behavioral objects of change contain psychological objects such as cognitive frames,
motivation and identity that are important for strategy formation and implementation
\citep{Vince2011,Gavetti2012} as they 'concern how decisions or actions are shaped,
subverted and transformed by emotions, and (..) determine the ”way we do things here”'
\cite[p. 338]{Vince2011}.

Moreover, these non-behavioral objects of
change are themselves far from static and are found to change over time as responses to
external change \citep{Zollo2016} thus making them a natural part of the mechanisms
underpinning how dynamic capabilities work \citep{Schilke2018,Verona2011}.  Still, far less
is know about the non-behavioral objects of change and how they work as a mechanism in the
dynamic capabilities theory.

We aim to contribute to bridging this gap by analyzing to what extent DCs influence
competitive advantage through two paths - one old, and one new - namely behavioral and
non-behavioral objects of change.  Specifically, we ask the following research question:

@@

To help answer this research question we build a conceptual model which illuminates the
relationships between $DC$, behavioral and non-behavioral objects of change and $CA$. The
model firmly integrates theorizing on organizational cognition \citep{Lakoff1987},
non-behavioral object of change \citep{Verona2011} and cognitive psychology
\citep{Tversky1983,Kahneman2011} into the $DC-CA$ nexus. Specifically, we focus on a
particular type of non-behavioral object of change, strategic cognition, and how it acts
as a mediator between dynamic capabilities and firm competitive advantage and hence has a
pivotal place in the literature. Reflecting this, we argue that DCs have two important
functions, namely 'DC as {\bf coach}' and 'DC as {\bf conductor}':

Dynamic capabilities take the function of a {\bf conductor} by orchestrating routines and
capabilities, and the role of a {\bf coach} in stimulating strategic cognition.  We discuss this
theoretical understanding of $DC$ and use the case of Microsoft’s transition into the cloud
computing business as an illuminating case. Importantly, this case demonstrates how the
roles of conductor and coach coexist and in generating strategic change and competitive
advantage.

The conceptual model is used to derive a set of hypotheses which is tested against
longitudinal survey data with a 10 year time lag. We find that $DC$ influence competitive
advantage through both behavioral and non-behavioral objects of change. Thus, the paper
empirically validates emergent theorizing proposing non-behavioral objects of change as a
new path between $DC$ and competitive advantage. Moreover, the paper helps to broaden the
conceptual basis for empirically investigating how $DC$ enable competitive advantage. This
has implications for understanding how dynamic capabilities work, as well as practical
implications for how strategic change is implemented in firms. Thus, the paper clarifies
the theoretical role of non-behavioral objects of change as well as a possible
operationalization through the concept of \emph{strategic cognition}. Moreover, we
conceptualize two different expressions of dynamic capabilities (i.e. conductor and coach)
and theorize on how they work in the $DC-CA$ nexus. Finally, we develop and test an empirical model
suited for analyzing mediating constructs on the dynamic capabilities-competitive
advantage nexus. The time-lag of our data in conjunction with advances in empirical
estimations of multiple simultaneous mediators makes for an interesting model to push our
understanding of important mechanisms in the theory.

@@

\section{Theoretical background and hypotheses}

Competitive advantage is considered the ’holy grail’ of strategic management
\citep{Schilke2018} and is indeed the focal point of important antecedent theories of
dynamic capabilities \citep{Arndt2018,Pezeshkan2016b}. Consequently, contributions to push
the theory forward should enable a better understanding of how dynamic capabilities work
to influence firms’ competitive advantage.

The collective term for assets available for firms in their pursuit of their objectives
are \emph{resources} \citep{Helfat2011,Helfat2007}. This term entails tangible assets such
as machines, capital, and labor, but also intangible assets such as procedures, practices,
and intellectual property \citep{Pisano2017}. The process of using these resources in the
pursuit of business objectives such as competitive advantage is a firm's
\emph{capabilities} which can be a set of routines and skills applied
\citep{Helfat2007,Helfat2011}. Resources and capabilities are both tied organizational
behavior and how the organization act in their operation. Hence, they are coined
\emph{behavioral objects of change} \citep{Verona2011}: 'A pattern of action, a process,
an operating routine, or any form of group activity characterized by some level of
stability and predictability' (ibid. p.538). Such behavioral objects such as capabilities
are the conventional understanding of how dynamic capabilities work; they 'purposefully
create, extend and modify the resource base' \cite[p. 4]{Helfat2007}. In the terminology
proposed in this paper this process takes the shape as a {\bf conductor} envisioning a
conductor orchestrating firm capabilities. Extant literature on dynamic capabilities are
increasingly focusing on how firm resources are changed by dynamic capabilities as a
response to changes in the environment \citep{Schilke2018,Eriksson2014}, thus placing firm
capabilities as a mediator between dynamic capabilities and competetive advantage
\citep{Helfat2007}. This notion of 'orchestration' of firm resources, assets and
capabilities has indeed been emphasized in the extant literature
\citep{Sirmon2011,Helfat2015}. Thus the notion of a {\bf conductor} captures a directional
role played by $DC$ in terms of its functioning to 'orchestrate' \emph{behavioral}
objects of changes such as \emph{operational capabilities}.

However, as \cite{Verona2011} points out: 'This approach misses the fundamental aspect
that organizational capabilities specific to managing change can hardly be reduced to the
management of behavioral change' (ibid p. 538). This opens for \emph{non-behavioral
  objects of change} which are defined as 'behavioral antecedents, such as cognition
(e.g. cognitive frames) and motivation' (ibid p. 539). This concept entails the 'how
decisions or actions are shaped, subverted, and/or transformed by emotions; and it
concerns how emotions become embedded in cultural and political practices that determine
the "way we do things here"' (Vince and Gabriel 2011 . ch 15). One particular dimension of
these non-behavioral objects of change particularly conducive to the context of
generating strategic change is the \emph{strategic cognition} of the
organization. Strategic cognition can be defined as conceptual structures in the minds of
the individuals that encapsulate a shared understanding of the reality the
individuals face with respect to the strategic goal of the organization \footnote{This
  definition is based on the definition of \cite{Lakoff1987} definition of the more generic
  'cognitive representation' and is adapted to fit the pursuit of strategic goals such as
  competetive advantage which is the aim of this paper}. This strategic
cognition is the result of a mental process that underlie internal
audiences adoption of new strategic representation and identity code to move toward a
new or existing goal \citep{Gavetti2012}. Although strategic cognition may lead to
behavioral outcome, the nature of the concept itself is tied to cognition rather than
actual latent action \citep{Zollo2016}. Thus, dynamic capabilities are not sufficiently
creating strategic change through its role as {\bf conductor}. Additionally they need to
exercise a role to adapt the strategic cognition of the organization to create
a shared understanding of the sense and necessity of the strategic goals of the
orientation. We coin {\bf coach} to capture the enabling role dynamic
capabilities plays through \emph{strategic cognition} as
juxtaposed to the more directional role played as {\bf conductor} of \emph{operational
  capabilities}. According to the Cambridge dictionary the noun 'coach' means 'someone
whose job is to teach people to improve' and 'someone whose job is to train and
organize'. Albeit often used in relation to sports it is also increasingly used for any
process of 'learning and development intervention that uses a collaborative, reflective, goal-focused
relationship to achieve professional outcomes' \cite[p. 253]{Jones2016}. We use the word
as an analogous idiom for an organizational process focusing on the subject (e.g. the firm or an
individual) to utilize own capacities as well as activating latent motivations. Thus it is
juxtaposed to a {\bf conductor} that creates results through directions and
orchestration.

\subsection{Conducting and coaching at Microsoft}\label{sec:case}

@@

The non-behavioral objects of action has no similar agreed upon analogy, so for the
purpose of our analysis we envision this mechanism to work more as a cognitive process
enabling the organization to tear away from their current understanding and ways in which
things are done \citep{Vince2011}. Whereas the analogy of a {\bf conductor} suggests a
function that composes an outcome based on the available capabilities, the notion of a
{\bf coach} refers to the function of enabling others to achieve their objectives. The
process is all about getting employees on board, and by changing an organization`s
thinking from the automated and well-known, to the more cognitively aware. We will now
turn to the two functions of dynamic capabilities in turn.

@@

The bulk of prior contributions study firm capabilities and resources as the mechanism
through which dynamic capabilities influence an outcome
\citep{Pezeshkan2016b,Fainshmidt2016a,Protogerou2012,Schilke2018}. Particularly, much of
the extant literature concerns measures of accounting profitability and competitive
advantage as the final outcome of dynamic capabilities
\cite{shamsie2009,Schilke2014,Schilke2014a,Teece2016}. Thus, from these prior
contributions we derive a fundamental hypotheses of the relationship between dynamic
capabilities and competitive advantage:

@@

vJuxtaposed to the experiences from Microsoft, the case of Polaroid brings another set of
experiences to our discussion on non-behavioral objects of action (see \citep{Tripsas2000}
for a detailed account). Cognitive inertia on the part of the organization played a
pivotal part in the failing of the company`s digital imaging effort \citep{Verona2011}. In
other words, Polaroid did not exercise dynamic capabilities to change the cognition of the
organization. Similar patterns are found in Kodak and Anderson Consulting
\citep{Kaplan2005}, and Smith Corona \citep{Danneels2010}. These examples highlight the
absence of 'dynamic capabilities specific to the adaptation of cognitive frames, first,
and consequently of operating processes' \citep[p. 541]{Verona2011}. This is also
highlighted at the core of the dynamic capabilities literature: 'Dynamic capabilities are
about doing the right things, at the right time, based on new product (and process)
development, unique managerial orchestration processes, a strong and {\bf change-oriented
  organizational culture}' (our emphasis) \cite[p]{Teece2014}.

When dynamic capabilities are executed it is as a response to changes in the environment
in order to stay competitive. This processes relates to creating strategic change
\citep{Helfat2007} and entails changing strategic direction and objectives of the
organization. This strategic change requires, in addition to modifying, extending and
creating firm capabilities, also some kind of alignment with the organization as a whole -
what is often referred to as 'how we do things here' \citep{Nelson1982}, and how an
organization think and act \citep{Nadella2017}. This alignment is possible only if the
employees share a common cognitive frame related to the the strategic direction of the
company enabling them to make sense of the changes ahead. This is what we have coined
\emph{strategic cognition} and has its expression in Microsoft's development of a growth
mindset. Strategic cognition is needed to get employees on board as well as capture the
efficiency of non-directed change. Getting on board refers to the general willingness of
the organization to accept and value the importance of some decision or system 'without
being prompted or required by formal governance
mechanisms'\citep[p. 812]{Simsek2009}. Moreover, When individuals of a firm buys into
changes they will enhance the full organization`s ability to not only act upon changes
\citep{Marchand2004}, but also in 'timely responsiveness' \citep{Kohli1993}.

Seminal contributions in cognitive psychology describes this state of actively engaging in
cognitive processes 'system 2 thinking' and define it as allocating 'attention to the
effortful mental activities that demand it, including complex computations'
\cite[p. 21]{Kahneman2011}. This is juxtaposed to the more automated 'system 1' which
'operates automatically and quickly, with little or no effort and no sense of voluntary
control' (ibid). The process is all about changing an organization`s thinking from the
automated and well-known, to the more cognitively aware. The extent to which individuals
in an organization is enabled to allocate attention to effortful mental activities related
to understanding needs for strategic change instead of relying on automatic reasoning
\citep{Tversky1983} or rules of thumb \citep{Kahneman1979}, can be a forceful remedy
aganist organizational inertia \citep{Adriaenssen2016}.

Consequently, strategic cognition not only contributes to getting people on board, it also
maintain a more efficient change process where the individuals take more responsibilities
for the change itself and leverages their individual knowledge in the endeavor. For a firm
to create real strategic change as enabled by their dynamic capabilities and hence stay
competitive, is will need to act as a {\bf coach} to instill a strategic cognition in the
organization.

@@

@@

We use a unique data set where Norwegian technology firms were administered a survey at
two points in time. ($T_0 = 2005$ and $T_2 = 2014$). The data combine the secondary data
from an earlier survey with registery data and a follow-up survey conducted by the
authors. The baseline survey was captured in 2005 when a web based survey was sent to 1721
technology firms in Norway intending to capture the concept of dynamic capabilities
\citep{Alsos2007}. Approximatly 70\% of the firms returned filled-in questionaires mouting
to a total sample size of 1199 firms. Details of the data gathering process and the
validation is available in \cite{Alsos2007}, but a brief explanation is waranted. The
authors used a mixed methods approach combining an extensive literature review with
qualitative interviews and testing in the field. Face validity of the items was further
examined by testing the items on experts. The authors of the present paper did take part
in this data initial collection. However, we have been given access to the raw data
collected in 2005. The second author of this paper followed up with a similar survey in
the spring/summer of 2014, 9 years after the first round. All the same firms received a
web-based questionnaire containing the same measures of the key constructs in this
paper. Following htis second round we were left with a sample size of 283 representing the
number of firms returning a filled-in second questionnaire.

The population was all businesses registered to a scheme for
tax deduction of R\&D costs (called SkatteFUNN). As all enterprises which are eligible for
taxation could register their R\&D activities to receive a tax refund, the registered
enterprises include close to all enterprises which are involved in such activities at the
time of our study. The context of R\&D active firms is particularly relevant for studying
dynamic capabilities because R\&D activity is a signal of responsiveness to change and
that the firms are located in industries with a certain level of dynamism.

@@

\section{Results}
The main results of our analysis are depicted both in the graphical path diagrams (figure
\ref{fig:res}) and the regression table in table \ref{tab:res}. The use of PLS-SEM has
increased substantially in the realm of marketing \citep{hult2018} and strategic management
\citep{Hair2012,Valaei2017} and is the topic of special issues in leading journals \citep{Hair2020} (e.g. Journal of
Business Research call for papers in 2020). In this tradition we aim to present thorough
analysis of our models and largely follows the recommendations made in recent
contributions \citep{Hair2019}.

@@

\section{Discussion}
The aim of this paper is to theorize about how non-behavioral objects of change work as a
mechanism through which dynamic capabilities influence competitive advantage. Our results
indicate that both behavioral (\emph{operating capabilities}) and non-behavioral
(\emph{strategic cognition}) mediates the effect from dynamic capabilities on competitive
advantage. The first mechanism, dubbed 'conducting' in reference to the well established
terminology of 'resource orchestration' \citep{Sirmon2011,Helfat2007}, shows a robust and
strong function as a mediator in line with our first hypothesis. When firms exhibit strong
dynamic capabilities they are able to adapt their operating capabilites to stay
competitive. We hypothesized that this relationship would be positive and significant. Our
results indicated exactly this relationship and thus lending support to hour hypothesis.
In our example context of Microsoft Nadella did several investments in capabilities and
resources as a response to Microsofts dynamic capabilities execution. For example, he
acquired important capabilities from M\&A efforts, and upgraded the competencies regarding
cloud business and big data \citep{Nadella2017}. This should not be surprising given the
vast literature on indirect effects of dynamic capabilities
\citep{Schilke2018,Eriksson2014} supported by empirical and theoretical research
\citep{Pezeshkan2016b,Fainshmidt2016a,Protogerou2012,Karimi2015}. However, it adds to this
ongoing effort of investigating mechanisms through which dynamic capabilities work which
is called for in the literature \citep{Schilke2018}.

### 2019-12-11 article4_tommylars.rnw
Ready for submission

@@

\emph{H2a: When dynamic capabilities are being mediated by VRIO resources the firm can
  achieve competitive advantage}

>>>>

Extant literature on dynamic capabilities are increasingly focusing on how firm resources
are changed by dynamic capabilities as a response to changes in the environment
\citep{Schilke2018,Eriksson2014}, thus placing firm resources as a mediator between
dynamic capabilities and competetive advantage \citep{Helfat2007}.

@@

\emph{H2a: When dynamic capabilities are being mediated by VRIO resources the firm can
o  achieve competitive advantage}
